---
title: "Statistical inference with the GSS data"
output: 
  html_document: 
    fig_height: 4
    highlight: pygments
    theme: spacelab
---

## Setup

### Load packages

```{r load-packages, message = FALSE}
library(ggplot2)
library(dplyr)
library(statsr)
```

### Load data

```{r load-data}
load("gss.Rdata")
```



* * *

## Part 1: Data
The GSS data has been through  interviews . The data has been collected through random sampling . As the survey is conducted by random sampling, the results from the survey or rather the inferences drawn from this survey can be generalized to the entire US population.

* * *

## Part 2: Research question
To check whether there is a association between reading newspapers and confidence level in goverment. SInce the data has a specific branch of government , i would like to check whether reading newpaper associates with people's confidence in the executive branch of federal government.

Newspapers are a part of the daily lives of almost all the people. Hence, testing whether reading newspapers affect people's confidence in gov. would be a good thing to explore .
 

* * *

## Part 3: Exploratory data analysis

Lets only talk about the 21st century data.
```{r}
summary(gss$year)
dataset<-gss%>%filter(year>=2000)

dataset$year%>%summary()

```

```{r}
dataset %>% select(news) %>% table()
```
It seems like majority people read newspaper everday which indeed is a good thing.
Lets see how much NA values does our concerned columns have.

NOTE:- NULL VALUES WOULD NOT EFFECT THIS CASE BUT CLEANING DATA IS GOOD PRACTICE IN MY ACCORDANCE.
```{r}
any(is.na(dataset))
library(Amelia)
colnames(dataset)
missmap(dataset[,c(52,105)],main = "yellow-missing, black-not missing",legend = F,col = c("yellow","black"),y.at = c(1),y.labels = c(''))

```
Since we are dealing with categorical variables, we cannot replace these values with mean . So we may remove these.

```{r}
df<-dataset %>% filter(!is.na(news)) %>% filter(!is.na(confed))
missmap(df[,c(52,105)],main = "yellow-missing, black-not missing",legend = F,col = c("yellow","black"),y.at = c(1),y.labels = c(''))
table(df$news)
```

Lets see how many people have confidence on the federal section of the government

```{r}
table(df$confed)
```
Hmmm.., that looks like most of the people donot have much confidence on this gov section.

Lets visualize the distribution among news and confidence.
```{r}
contin_table=table(df$news,df$confed)
mosaicplot(contin_table,shade = T)
```
It seems that people who read news papers don't trust the gov.'s section.

* * *

## Part 4: Inference

Since I are dealing with multiple factors and I want to test independence, I would use the chi-square test. 

My null hypthesis would be- There is no relation between reading newspapers and having confidence on the government's federal section.

Alternale hypothesis - There is indeed a  relation between reading newspapers and having confidence on the government's federal section.

```{r}
chisq.test(x = df$news,y = df$confed)$expected
```
Observed:-

```{r}
table(df$news,df$confed)
```

Conductiong the test:-

```{r}
chisq.test(x = df$news,y = df$confed)
```

The obtained test statistic is 30.921 with a p-value 0.0001451. Since, p<0.05, we would reject the null- hypothesis .

Hence, there indeed is a relationship between reading newspapers and having confidence on the federal section of the government.













